Prelim Outline

### 2019-07-26 outline.tex
worked on propsed outline. ready for writing, has TBLG stuff now

\section{Introduction and methods}
\paragraph{Overall goal} We want to be able to systematically develop low-energy effective models for strongly correlated electron systems.

\paragraph{Barrier to achieving goal} Non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge - makes writing down low-energy effective theories for strongly correlated materials very challenging.

\paragraph{Current state of the art} \textcolor{red}{Talk about methods people use for model fitting now. Will fill out as I do literature review.}

\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.

\paragraph{Density matrix down folding} 
The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with a higher energy Hamiltonian, like the \textit{ab-initio} Hamiltonian $H_{ab}$.

\paragraph{Fixed node diffusion Monte Carlo}
We will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate \textit{ab-initio} energies and reduced density matrices.

@@

\section{Proposed work}
\paragraph{Hubbard: Effective models at half-filling} We want to develop low-energy effective theories for the half-filled Hubbard model over the space U: 0 - 12, 1/$k_B$T: 0 - 3 using the DMD method, particularly on the medium coupling region of U: 2 - 6 and 1/$k_B$T $\sim$ 1 where perturbative effective theories break down and a metal-insulator transition exists.

\paragraph{Hubbard: Effective models away from half-filling} A farther reaching goal would be to develop effective theories in the U-T-$\delta$ space using DMD, where more complex behavior similar to the cuprates have been found.

\paragraph{Hubbard: Finite size effects} Our initial calculations necessary for DMD can be done using exact diagonalization, but memory limitations will require us to use approximate methods like lattice quantum Monte Carlo to calculate energies and density matrix elements for states in our LES.

\section{Proposed timeline}
\textit{Do this after proposed work nailed down}.

>>>>

\section{Introduction}
\begin{itemize}
\item Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.

\item The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.

\item Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).

\item We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
\end{itemize}

\section{Methods}
\begin{enumerate}
\item Density matrix downfolding
\begin{itemize}
\item We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.

\item The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto an equivalent linear regression problem.

\item This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.

\item Generally, one cannot sample the true low-energy subspace and an approximate subspace $\mathcal{LE}^\prime$ is sampled.

\item Next, a set of candidate descriptors which form the effective Hamiltonian are selected.

\item Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.

\item Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.

\item Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.
\end{itemize}

\item Fixed-node diffusion Monte Carlo 
\begin{itemize}
\item Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.

\item This stochastic implementation of DMC suffers from a fermion sign problem which is alleviated via a fixed-node approximation.

\item We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
\end{itemize}
\end{enumerate}

@@

\section{Proposed work}
\begin{itemize}
\item My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.

\item The first system I will develop a model for is the benzene molecule.

\item While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.

\item As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.

\item The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.

\item A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.

\item Note also that a model for benzene has already been developed using DMD.

\item This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.

\item The second system I will be interested in working on is SLG.

\item The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.

\item I will extend the previous model by introducing long range density-density interactions into the effective theory.

\item The candidate model I anticipate to work with is an extended Hubbard model of the form.

\item CVMC will be used to sample the low-energy wave functions necessary to fit the model.

\item The last systems I will develop models for are AA and AB stacked BLG.

\item Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.

\item I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.

\item I will again use CVMC to generate the low-energy wave functions for the model regression.

\item The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.

\item I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.

\end{itemize}
\section{Proposed timeline}
\textit{Do this after proposed work nailed down}.
